Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-06-02 doc/paper.tex
ALMOST finished with Results

\begin{abstract}
An increased focus on a
S \zh{我把} O V
is proposed.

A particular system for doing this is presented.
It allows syntactic complexity via rules.
It also allows high-quality examples via templates, with minor syntactic modifications via rules.

Preliminary results do not show much improvement.
This is because we need to scale up?

Future work is proposed.
In particular, we believe this is amenable to crowdsourcing by ``amateur linguists''.

@@

Large, high quality synthetic parallel corpora could be useful as:

@@

Our parallel corpus generation system (Section \ref{sec:generator}) is very similar to the recursive equivalence class approach to example-based MT \mycitep{ebmtbook} used by \mynewcite{brown:99}. 
Our generator has more of a rule-based flavor, using a somewhat more varied grammar that also admits the use of participial noun modifiers.
But more importantly, we believe this line of work was ahead of its time and consider our main contribution to be bringing attention to its potential value to the contemporary machine learning approach to MT.

@@

% TODO: looks like this is GOING to run over length. save some of this for AMTA, which is going to run way UNDER length
%\subsection{Simplistic Linguistics}
We adopt an exceedingly simplistic view of language as nouns, verbs, and their modifications and elaborations.
Inspired by early work on transformations by \mynewcite{chomsky}, we focus on declarative indicative present active sentences in the hopes that they can be transformed into other forms in the long run. 
Similarly, although we enforce some semantic constraints, our system generally produces nonsensical sentences like ``{\em Colorless green ideas sleep furiously.}''
The fact that verb phrases can also serve as nouns (gerundives) and modifiers (participles) appears to be a common source of complexity, and we made it a priority to implement participial modifiers.
To capture recursive structures like participles, we adopt a rudimentary tree representation that is something intermediate between constituency trees and dependency trees, with constituent-like modifiers that also maintain dependency-like links to their targets.

@@

% TODO: should I bother with this? 
Crude though it may be, we plan on eventually open-sourcing both our generator and its accompanying data

@@

@@

\subsection{Methods}

% TODO: filler for AMTA? used for corpus analysis stanford POS tagger (incomplete bibtex entry?) \mycitep{postagger:2003}

As mentioned in 
The IWSLT 2015 version of the WIT$^3$ corpus, which consists of multilingual subtitles of TED talks, covering a wide variety of subjects. 
considered briefly in Section \ref{subsec:case_study}

As prepared lectures, they fall somewhere between the conversational language of movie subtitles and the formal written language of UN resolutions.
The latter are freely available online from the OPUS project \mycitep{opus:2012} as the OpenSubtitles\footnote{http://www.opensubtitles.org/} 2016 \mycitep{opensubtitles:2016} and the MultiUN \mycitep{multiun} corpora, respectively.

We consider the Chinese to English machine translation task, using the

first

Out of domain corpus

we of 
the {\small \tt tmcombine} script \mycitep{tmcombine} included with Moses

the latest version of the Stanford Segmenter \mycitep{segmenter:2005,segmenter:2008}

citations needed
IWSLT \mycitep{wit3} - point out that this corpus is freely available
OPUS 
zh\_zh corpus from the OpenSubtitles 2016 corpus , with lines cleaned using some ad hoc heuristics
MultiUN 
Moses \mycitep{moses}

Since we are interested in the relative effect an additional synthetic parallel corpus
we choose a simple baseline system and fix other variables

We configured Moses as described in \mycitep{wit3}, with the exception of changing the target monolingual language model toolkit to KenLM \mycitep{kenlm}, which we found to be faster and stabler than IRSTLM \mycitep{irstlm} at decoding time.
For the remaining unspecified model settings, we used the {\small \tt grow-diag} alignment heuristic recommended by \mynewcite{segmenter:2008} and the {\small \tt mslr-fe} configuration for lexicalized reordering.

The corpus, which is freely available, consists

IWSLT: TED talks are a series of lectures, which makes them

heuristically purged zh\_zh, which is a very noisy corpus

\subsection{Results}

2016-02-20 learning curves adding only 130,000 in-domain lines improves scores by like 2.0 BLEU points

2016-03-03-225556 1M zh\_zh lines made scores slightly WORSE?

2016-04-29-093202 1M MultiUN lines only improved scores SLIGHTLY (by like 0.4 BLEU points)

Quantity and maybe even quality (you'd expect MultiUN translations to be better?) matters much less than having data in-domain

Of course, I need to redo (ALL?) these jobs with final segmenter choice

Evaluated using the IWSLT 2015 progress data set (tst2014) to allow comparison with both 2015 and 2014 systems.

training curves (just to prove my controls aren't wrong)
my synthetic corpora don't help. 
but in-domain corpus helps significantly more than out-of-domain. (this is a commonly reported result)

Bottom line
My corpora don't really seem to help, but neither do out-of-domain corpora.

this will easily take up a whole column, because it includes 2 tables...

>>>>

\begin{abstract}
The large-scale generation of high-quality synthetic parallel corpora for data-driven machine translation is proposed.

A preliminary system for doing this is presented,

syntactic complexity in the spirit of rule-based machine translation,

as well as high-quality sentence pairs via templates, in the spirit of example-based machine translation.

Early results do not show much improvement.
This is because we need to scale up?

Future work is proposed, in particular, 
In particular, we believe this is amenable to crowdsourcing by ``amateur linguists.''

@@

Large, high-quality synthetic parallel corpora could be useful as:

@@

Our parallel corpus generation system (Section \ref{sec:generator}) is very similar to the recursive equivalence class approach to example-based MT \mycitep{ebmtbook} used by \mynewcite{brown:99}. 
Our generator has more of a rule-based flavor, using a somewhat more varied grammar that also admits the use of participial noun modifiers.
But more importantly, we believe this line of work was ahead of its time and consider our main contribution to be bringing attention to its potential value to contemporary MT.

@@

% TODO: looks like this is GOING to run over length. save some if not all of this for AMTA, which is going to run way UNDER length
%\subsection{Simplistic Linguistics}
We adopt an exceedingly simplistic view of language as nouns, verbs, and their modifications and elaborations.
Inspired by early work on transformations by \mynewcite{chomsky}, we focus on declarative indicative present active sentences in the hopes that they can be transformed into other forms in the long run. 
Similarly, although we enforce some semantic constraints, our system generally produces nonsensical sentences like ``{\em Colorless green ideas sleep furiously.}''
The fact that verb phrases can also serve as nouns (gerundives) and modifiers (participles) appears to be a common source of complexity, and we made it a priority to implement participial modifiers.
To capture recursive structures like participles, we adopt a rudimentary tree representation that is something intermediate between constituency trees and dependency trees, with constituent-like modifiers that also maintain dependency-like links to their targets.

@@

% TODO: should I bother with this? 
Crude though it may be, we plan on eventually open-sourcing both our generator and its accompanying data.

@@

% running over length, but let's just write it all down and then see what to cut (hint: that stupid tst2010 stuff. maybe just summarize it in a footnote)

@@

%\subsection{Methods} % no space? just do one paragraph each?
We consider training set amplification as a first application.
Since we are interested in the relative effect of our synthetic parallel corpus, we chose a simple baseline system and fix all other variables.
Namely, we used the baseline configuration of Moses \mycitep{moses} as described in \mycitep{iwslt2015}, with the exception of changing the target language model to KenLM \mycitep{kenlm}, which we found to be faster and stabler than IRSTLM \mycitep{irstlm} at decoding time.
For the remaining unspecified model settings, we used the {\small \tt grow-diag} alignment heuristic recommended by \mynewcite{segmenter:2008} and the {\small \tt mslr-fe} configuration for lexicalized reordering. 
Chinese text was preprocessed using the latest version of the Stanford Segmenter \mycitep{segmenter:2005,segmenter:2008}.
Phrase and lexicalized reordering tables were interpolated using the {\small \tt tmcombine} script included with Moses \mycitep{tmcombine}, with the interpolation tuned on the {\small \tt dev2010} data set.
BLEU scores were computed as described by \mynewcite{iwslt2015} using the IWSLT 2015 progress data set ({\small \tt tst2014}).\footnote{We chose the {\small \tt tst2014} test set to facilitate comparison with systems from both IWSLT 2014 and 2015 in future work.}

%\subsection{Task}

% TODO: filler for AMTA? used for corpus analysis stanford POS tagger (incomplete bibtex entry?) \mycitep{postagger:2003}
We evaluate the contribution of our synthetic parallel corpora using the Chinese to English task of IWSLT 2015 \mycitep{iwslt2015}, mentioned briefly in Section \ref{subsec:case_study}.
This task uses the WIT$^3$ corpus \mycitep{wit3}, which consists of multilingual subtitles of TED talks covering a wide variety of subjects. 
As a control, we also add the freely available MultiUN corpus \mycitep{multiun} from the OPUS project \mycitep{opus:2012}.

% TODO: add OpenSubtitles 2016 results? they don't really add much to the discussion - I can just mention them briefly, maybe even in a footnote
% oh okay, this is nice - if I don't include OpenSubtitles, then I don't have to talk about the stupid purging either
%heuristically purged zh\_zh, which is a very noisy corpus
%zh\_zh corpus from the OpenSubtitles 2016 corpus , with lines cleaned using some ad hoc heuristics

%\subsection{Results}

% yeah sure, let's add Bing/Google results just for kicks, then take them out later

% definitely want 2 tables?
    % one for learning curves, to show that iwslt > zh_zh > multiun - domain overlap is greater

    % one for 100%, to show that the out of domain corpora aren't helping that much. need moar in-domain data (probably)
    % well, maybe i'll do away with this one

    % I guess don't bother showing single-corpus results for zh_zh and multiun, because then I'd have to show MY single-corpus results, which are even more embarrassing than the horribleness I've written up here

% when i actually go and READ the table, it's clear that I don't even want to LOOK at more than one metric. it's information overload.
%\begin{table}
%\centering
%\label{table:tst2010}
%\begin{tabular}{|l|r|r|r|}
%\hline
%                    & BLEU  & NIST & TER    \\ \hline
%IWSLT15 official    & 11.43 & 4.67 & 72.65  \\ \hline
%Our baseline        & 11.54 & 4.43 & 72.43  \\ \hline
%  + OpenSubtitles   & 11.54 & 4.43 & 72.43  \\ \hline
%  + MultiUN         & 11.67 & 4.48 & 72.60  \\ \hline
%  + generated       & 11.46 & 4.39 & 72.51  \\ \hline
%% &  &   \\ \hline
%\end{tabular}
%\caption{}
% 
%\end{table}

% side-by-side tables as in emnlp2016.tex
\begin{table}
\label{tab:results}
\centering
\small
\begin{tabular}{cc}

\begin{tabular}{|l|r|}
\hline
                    & BLEU   \\ \hline
25\% corpus         &  9.95  \\ \hline
  + 75\%            & \textbf{11.47}  \\ \hline
%  + Subtitles       & 10.57  \\ \hline
  + MultiUN         & 10.76  \\ \hline
  + Generated       &  9.91  \\ \hline
\end{tabular}

\begin{tabular}{|l|r|}
\hline                    & BLEU   \\ \hline
Official            & 11.43  \\ \hline
Our baseline        & 11.54  \\ \hline
%  + Subtitles       & 11.54  \\ \hline
  + MultiUN         & \textbf{11.67}  \\ \hline
  + Generated       & 11.46  \\ \hline
\end{tabular} & % oh this is how you get 2 tables side by side

\end{tabular} % cc (i.e., the big super-table)

\caption{
    (Left) Chinese-to-English BLEU scores for our baseline Moses configuration with phrase and reordering tables trained on a random 25\% subset of the IWSLT15 training corpus and interpolated (+) with tables for the remaining 75\% (\textasciitilde 150K) of the in-domain corpus, the MultiUN corpus (a \textasciitilde 1M line subset), and a synthetic corpus from our generator (\textasciitilde 5M lines).
    (Right) Same as on the left, but starting with tables for the entire IWSLT15 corpus.
    Official published scores of the IWSLT15 baseline system are also given for comparison.    
    Qualitatively similar trends for NIST and TER scores were also observed (omitted for clarity).
%    The ``baseline'' row gives 
%    MultiUN is actually more data because it has more words per line.
% automatically cleaned version of the en-zh_zh language pair of OpenSubtitles 2016, roughly 1M lines.
% and a similar number of lines from MultiUN, which is actually more words. but it doesn't do as well.
} % caption

\end{table}

% preemptively answering reviewer question: yeah yeah, the trained tmcombine weights are like 0.001 for my generated corpora
The results of these experiments are shown in Table \ref{tab:results}.
Our generated corpora do not make a significant impact on the final BLEU score,\footnote{
    Results for the most recently generated synthetic corpus are shown; scores slightly higher than our baseline have been obtained on older generated corpora.
    We believe both to be within the error bars of the MERT tuning procedure.
}
indicating that our generator has a long way to go.

%%%%%%%%%%%%%%%%%%%%%
TODOOOOOOOOOOOO
% TODO: HOW SO?
wanted universal translation memory
much less ambitious goal of training set amplification
- out of domain data only boosts performance slightly
phrase table interpolation using the tmcombine script included with Moses
The results show that our generator has a long way to go.
Quantity and maybe even quality (you'd expect MultiUN translations to be better?) seems to matter much less than whether the additional data is in-domain
So lots of different basic syntactic variation didn't help
vocabulary is much too small? (only about 100 nouns currently. we used Stanford POS tagger \mycitep{postagger:2003} to find most common nouns, since different corpora tend to translate words differently, and the mini-goal here was training set amplification)
more templates
work on that is currently in progress. % I have never liked the sound of ``currently underway''
%%%%%%%%%%%%%%%%%%%%%%%%

For the table on the left, note that adding in-domain data improves performance significantly more than a much larger amount of out-of-domain data (MultiUN), a commonly reported phenomenon \mycitep{tmcombine}.
As prepared lectures, TED talks fall somewhere between the conversational language of movie subtitles and the formal written language of UN resolutions; data from the OpenSubtitles 2016 corpus \mycitep{opensubtitles:2016}\footnote{http://www.opensubtitles.org/} from the OPUS project are similarly ineffective at improving performance (scores not shown).
For the table on the right, note also that the same out-of-domain corpus now gives a much more modest performance increase.
These observations suggest that amplification of in-domain corpora should still be beneficial, particularly for low-resource language pairs or domains.

% no, I can't mention this, because it's just saying, ``duh, hope it's in-domain!''
%For a scaled-up corpus generator, one anticipates that the main concern would be to avoid skewing the phrase table with the ``unnatural'' synthetic corpus.

% TODO: maybe add Bing and Google results, since nobody seems to report these? save this for AMTA? (haha not gonna have room, sucka)
